Estimating contemporary migration numbers of adults for iteroparous species based on kinship relationships found between populations

### 2022-03-04 latex/ms.tex
reflect revision

%KOKOMADE enago%%%
\subsection{PO pair-based model}

Next, we consider the probability that offspring sampled in the population 1 and a parent sampled in the population 2 share an PO relationship, denoted by $\pi_{\rm PO, bet}$. $\pi_{\rm PO, bet}$ can be partitioned into the three probabilities: (i) the probability that sampled offspring in the population 1 born to a migrant; (ii) the probability that a migrant is sampled; and (iii) the probability that the migrant of a sampled offspring in the population 1 and the migrant sampled in the population 2 are identical. The first and third probabilities are the same as the one introduced in the previous subsection, that is, $2M/N_{1}$ and $1/M$. The second probability is $M/N_{2}$ by definition. Taken together, we obtain
\begin{align}
\pi_{\rm PO,bet} &= \frac{2M}{N_{1}} \frac{M}{N_{2}} \frac{1}{M} \nonumber\\
&= \frac{2M}{N_{1}N_{2}}.
\label{PO-1}
\end{align}
This form also holds under a flexible setting for the reproductive potential, as noted in {\bf Appendix}. If all $N_1$ parents safely move to the population 2 (i.e., $M = N_1$), $\pi_{\rm PO,bet}$ equals $2/N_2$; meanwhile, if the population 2 consists only of migrants (i.e., $M = N_2$), $\pi_{\rm PO,bet}$ equals $2/N_1$. These probabilities in the extreme cases are similar forms to PO probability that is randomly sampled within a population \cite[]{bravington2016close}.

Let $H_{\rm PO,bet}$ be the number of PO pairs found in offspring samples of size $n_{\rm O, 1}$ and parent samples of size $n_{\rm P, 2}$. Assuming that the total number of PO pairs in the two populations is much higher than $H_{\rm PO,bet}$, the distribution is approximated by a binomial form (i.e., $H_{\rm PO,bet} \sim {\rm Binom}[\pi_{\rm PO,bet}, n_{\rm O, 1} n_{\rm P, 2}]$); thus, the theoretical expectation of $H_{\rm PO,bet}$ is
\begin{align}
\mathbb{E}[H_{\rm PO, bet}] &= \pi_{\rm PO,bet} n_{\rm O, 1} n_{\rm P, 2} \nonumber\\
&= \frac{ 2 n_{\rm O, 1} n_{\rm P, 2} M } { N_{1} N_{2} } \label{PO-2-1} \\
&= \frac{ 2 n_{\rm O, 1} n_{\rm P, 2} m } { N_{2} }.
\label{PO-2-2}
\end{align}
The observed number of PO pairs in a sample is defined by $\widetilde{H}_{\rm PO, bet}$, and $\mathbb{E}[H_{\rm PO, bet}]$ in Equations \ref{PO-2-1} and \ref{PO-2-2} are replaced by $\widetilde{H}_{\rm PO, bet}$, generating the linear estimator of $M$ and $m$:
\begin{align}
\widehat{M_2} &= \frac{N_{1} N_{2} \widetilde{H}_{\rm PO,bet} } {2 n_{\rm O, 1} n_{\rm P, 2} },
\label{PO-3}
\end{align}
and
\begin{align}
\widehat{m_2} &= \frac{N_{2} \widetilde{H}_{\rm PO,bet} } {2 n_{\rm O, 1} n_{\rm P, 2} }.
\label{PO-4}
\end{align}

@@

Proposed estimators are based on the observed number of kinship pairs. Their expected number is linearly determined by the number of pairwise comparison (Eqs.~\ref{HS-2-1}-\ref{HS-2-2} and \ref{PO-2-1}-\ref{PO-2-2}), providing guidance for a sample size to ensure the condition that at least one or more kinship pairs can be found. The conditions are given by
\begin{align}
n_{\rm O, 1} n_{\rm O, 2} &> \frac{ N_{2} } { 4m },
\label{RS-1}
\end{align}
and
\begin{align}
n_{\rm O, 1} n_{\rm P, 2} &> \frac{ N_{2} } { 2m }.
\label{RS-2}
\end{align}
Roughly speaking, required sample size to ensure above conditions is that $n>\sqrt{N_2}$, where $n = n_{\rm O, 1}=n_{\rm O, 2}=n_{\rm P, 2}$. It should be noted that very tiny $m$, which dramatically increases the required sample size, produces significant genetic differences between two populations. In such a case, there are several methods to estimate migration rate by population genetics technique; alternatively, we focus the situation of moderate/large migration rate, which leads little genetic differences between two populations.

Other guidance for the required sample size is obtained by an approximate lower bound on the CV (coefficient of variation) of $M_1$ or $m_1$, defined by $1/\sqrt{H_{\rm HS, bet}}$, which is applied in the context of a classic mark-recapture \cite[]{seber1982estimation} or close-kin mark-recapture \cite[]{bravington2016close}. For example, to achieve a 30\% CV, the target of $H_{\rm HS, bet}$ is more than 10, providing the required sample size when population parameters are given (e.g., $N_1$, $N_2$ and $M$). Similarly, the CV of $M_2$ or $m_2$, defined by $1/\sqrt{H_{\rm PO, bet}}$, provides the required sample size for estimating $M_2$ or $m_2$.

@@

When the both HS pairs and PO pairs, which are found between populations, are simultaneously available, we can obtain efficient estimators by combined $\widehat{M_1}$ and $\widehat{M_2}$ for migration number: 
\begin{align}
\widehat{M_3} &= \frac{ N_{1} N_{2} \left(\widetilde{H}_{\rm HS,bet} + \widetilde{H}_{\rm PO,bet}\right) } {2n_{\rm O, 1} \left(2 n_{\rm O, 2} + n_{\rm P, 2}\right) },
\label{AP-1}
\end{align}
and by combining $\widehat{m_1}$ and $\widehat{m_2}$ for migration rate: 
\begin{align}
\widehat{m_3} &= \frac{ N_{2} \left(\widetilde{H}_{\rm HS,bet} + \widetilde{H}_{\rm PO,bet}\right) } {2n_{\rm O, 1} \left(2 n_{\rm O, 2} + n_{\rm P, 2}\right)}.
\label{AP-2}
\end{align}
For those estimators, $\widetilde{H}_{\rm HS,bet}$ and $\widetilde{H}_{\rm PO,bet}$ is weighted by a sample size $n_{\rm O, 2}$ and $n_{\rm P, 2}$, respectively.

\subsection{Estimation of parent number by PO pairs found within a population}

So far, we provide the formulation of the estimators, presented in Equations \ref{HS-3}-\ref{HS-4} and \ref{PO-3}-\ref{PO-4}, is a function with parent numbers for each population ($N_1$ and/or $N_2$); in other words, such estimators are available only when those parent number(s) are known. In this subsection, we provide the treatment for estimating the unknown parent numbers by additional usage of PO pairs, producing estimators for migrant number and migration rate that can be obtained only from the genetic data.

When PO pairs found in offspring samples and parent samples both from the population 2 in the second year are available, standard estimator of parental number can be obtained \cite[]{bravington2016close}, given by

\begin{align}
\widehat{N_2} &= \frac{ 2 n_{\rm O, 2} n_{\rm P, 2}  }{\widetilde{H}_{\rm PO, 2} + 1} ,
\label{AP-3}
\end{align}
where $\widetilde{H}_{\rm PO, 2}$ is the observed number of PO pairs found in offspring samples of size $n_{\rm O, 2}$ and parent samples of size $n_{\rm P, 2}$. The term ``$+1$'' reduces the bias especially when $\widetilde{H}_{\rm PO, 2}$ is small \cite[e.g.,][]{ecolevol2021p}, where a similar derivation of this bias correction is provided in \cite{Akita_2019}. By replacing $N_2$ by $\widehat{N_2}$ in Eqs.~\ref{AP-1} and \ref{AP-2}, we can obtain the following estimators:

\begin{align}
\widehat{M_4} &=  \frac{ n_{\rm O, 2} n_{\rm P, 2} N_{1} \left(\widetilde{H}_{\rm HS,bet} + \widetilde{H}_{\rm PO,bet}\right) } {n_{\rm O, 1} \left(2 n_{\rm O, 2} + n_{\rm P, 2}\right) \left(\widetilde{H}_{\rm PO, 2} + 1\right)},
\label{AP-4}
\end{align}
and
\begin{align}
\widehat{m_4} &=  \frac{ n_{\rm O, 2} n_{\rm P, 2}\left(\widetilde{H}_{\rm HS,bet} + \widetilde{H}_{\rm PO,bet}\right)} {n_{\rm O, 1} \left(2 n_{\rm O, 2} + n_{\rm P, 2}\right) \left(\widetilde{H}_{\rm PO, 2} + 1\right)}.
\label{AP-5}
\end{align}
While estimator of the migrate number (Equation~\ref{AP-4}) requires the (unknown) parent number in the population 1 ($N_{1}$), estimator of the migration rate (Equation~\ref{AP-5}) can be obtained only the observed kinship pairs.

Similar to estimating $N_2$, when PO pairs found in offspring samples and parent samples both from the population 1 in the first year are also available, we can obtain the estimator of $N_1$, written by
\begin{align}
\widehat{N_1} &= \frac{ 2 n_{\rm O, 1} n_{\rm P, 1} }{\widetilde{H}_{\rm PO, 1} +1},
\label{AP-6}
\end{align}
where $\widetilde{H}_{\rm PO, 1}$ is the observed number of PO pairs found in offspring samples of size $n_{\rm O, 1}$ and parent samples of size $n_{\rm P, 1}$. By replacing $N_1$ by $\widehat{N_1}$ in Eq.~\ref{AP-4}, we can obtain the estimator of migration number that can be calculated only from genetic data, given by
\begin{align}
\widehat{M_5} &=  \frac{ 2 n_{\rm O, 1} n_{\rm P, 1} n_{\rm O, 2} n_{\rm P, 2} \left(\widetilde{H}_{\rm HS,bet} + \widetilde{H}_{\rm PO,bet}\right)} {n_{\rm O, 1} \left(2 n_{\rm O, 2} + n_{\rm P, 2}\right) \left(\widetilde{H}_{\rm PO, 1} + 1\right) \left(\widetilde{H}_{\rm PO, 2} + 1\right)}.
\label{AP-7}
\end{align}

Table \ref{parameter} summarizes the conditions for estimators appeared in this paper as to whether $N_1$ and/or $N_2$ are known and which type of samples are needed for estimation, and provides required kinship type for calculating the estimator.

@@

The proposed estimators are derived under the assumption that reproductive potential is the same among individuals. As noted in {\bf Appendix 2}, this assumption can be relaxed and the estimators still hold, which is exemplified in iteroparous species that may show significant variation in reproductive potential among individuals. Furthermore, the estimators even hold for most of the case where mean reproductive potential are different between population 1 and 2, which is exemplified in the situation that the environmental condition for reproductive success varies between populations. The necessary condition is that the migration event does not depends on the degree of reproductive potential (see details in  {\bf Appendix 3}).

@@

We developed an individual-based model that tracks kinship relationships to evaluate the estimator's performance. The population structure was assumed to be identical to that in the development of the estimators. The population 1 and 2 comprised $N_1$ and $N_2$ parents with an equal sex ratio, and their offspring number was assumed to follow the geometric distribution with mean ${\bar \lambda_1}$ and ${\bar \lambda_2}$ (i.e., Poisson reproduction with mean $\lambda_{i,1}$ and $\lambda_{j,2}$ which follows the exponential distribution with mean ${\bar \lambda_1}$ and ${\bar \lambda_2}$), respectively. Migrant parents were randomly chosen from the population 1 at the end of the first year. Each offspring retained the parent's ID, making it possible to trace an HS and PO relationship.

Let a parameter set ($N_1$, $N_2$, $M$, ${\bar \lambda_1}$, ${\bar \lambda_2}$, $n_{\rm O,1}$, $n_{\rm O,2}$, $n_{\rm P,1}$, $n_{\rm P,2}$) be given. We simulated a population history and a sampling process, which generates proposed estimators; this process was repeated 1000 times, making it possible to construct the distribution of the estimators for each parameter set. All scripts and documentation for these analyses are available at \\https://github.com/teTUNAakita/CKMRmig.

@@

We numerically evaluated the performance of ${\hat M}$s or ${\hat m}$s for the case in which variable reproductive potential among parents. Scaled statistical properties of ${\hat m_1}$,  ${\hat m_2}$,  ${\hat m_3}$ and ${\hat m_4}$ are completely the same to ${\hat M_1}$, ${\hat M_2}$, ${\hat M_3}$ and ${\hat M_4}$, hereafter, we only demonstrate the results of ${\hat M_1}$-${\hat M_5}$. {\bf Figure~\ref{violins}} illustrates the distribution of the relative bias of ${\hat M}$ for limiting cases where parent and offspring sample numbers are identical (i.e., $n=n_{\rm P,1}=n_{\rm P,2}=n_{\rm O,1}= n_{\rm O,2}$) and parent numbers in the two populations are also identical (i.e., $N=N_1=N_2$). Relative bias is calculated by applying outputs of the individual-based model, which is defined as follows: ``(averaged estimator - true value)/true value.'' Full list of parameter sets to evaluate the performance (relative bias and CV of ${\hat M}$s) is summarized in {\bf Table S1} in {\bf Supporting Information}.

@@

First, we evaluated the accuracy of ${\hat M}$s based on their relative bias. As expected, for most of the investigated parameter sets, we observed that their relative bias is less than 5\%, as demonstrated in {\bf Table S1} in {\bf Supporting Information}. Thus, when the assumption that kinships are detected without any error is satisfied, it is reasonable to call ${\hat M}$s (nearly) unbiased estimators.

Next, we evaluated the precision of ${\hat M}$s based on their CV. {\bf Table S1} in {\bf Supporting Information} demonstrated the value of the CV; meanwhile, the violin plot ({\bf Figure~\ref{violins}}) visually provides the degree of precision. For each estimator ${\hat M}$, the precision increases with increasing of sample size. It should be noted that total sample size depends on the estimator. For example, $n=50$ in Fig.~\ref{violins}, the total sample size of ${\hat M_1}$, ${\hat M_2}$, ${\hat M_3}$, ${\hat M_4}$ and ${\hat M_5}$ is 100 ($=n_{\rm O,1}+n_{\rm O,2}$), 100 ($=n_{\rm O,1}+n_{\rm P,2}$), 150 ($=n_{\rm O,1}+n_{\rm O,2}+n_{\rm P,2}$), 150 ($=n_{\rm O,1}+n_{\rm O,2}+n_{\rm P,2}$) and 200 ($=n_{\rm O,1}+n_{\rm O,2}+n_{\rm P,1}+n_{\rm P,2}$), respectively.

As sample size or migrants increase, the precision increases and the shape of the distribution asymptotically becomes symmetric ({\bf Figure~\ref{violins}}). This is because an increasing of sample size or migrants is likely to increase the observed number of kinship pairs found between two populations ($\widetilde{H}_{\rm HS,bet}$ or $\widetilde{H}_{\rm PO,bet}$) and decreases the variance of those kinship pair numbers. In addition, the number of PO pair within a population ($\widetilde{H}_{\rm PO,1}$ or $\widetilde{H}_{\rm PO,2}$) also contributes the precision for estimating $N_1$ or $N_1$, respectively, providing ${\hat M_4}$ and ${\hat M_5}$ with a higher precision.

Our simulator can handle the situation with or without invasive sampling. Invasive sampling potentially affects the level of $m$ since sampled parents in the population 1 have no chance to move to the population 2 and thus those parents cannot be sampled. In this case, $m$ should be defined by $M/(N_1-n_{P,1})$ not by $M/N_1$, although it does not affect the estimator ${\hat M}$s, as demonstrated in {\bf Table S1} ({\bf Supporting Information}).

Finally, we investigated the situation in which the parental number in the population 1 is much more than in the population 2 (i.e., $N_1 \gg N_2$) and thus migrants are large proportion of $N_2$. In such a setting, e.g., $(N_1,N_2,M)=(10^4,10^3,500)$, we confirmed the similar property of ${\hat M}$s to the situation $N_1 = N_2$, as demonstrated in {\bf Table S1} ({\bf Supporting Information}), suggesting the robustness of ${\hat M}$s in heterogeneous population sizes.

@@

We theoretically developed estimators of the contemporary migration number (${\hat M_1}$--${\hat M_5}$) and migration rate (${\hat m_1}$--${\hat m_4}$) of parents between two pre-defined populations in iteroparous species. The proposed estimators are based on known PO relationship and HS relationships observed between and within two populations without any error of kinship assignment. Users can choose the estimator developed for the situation as to whether the parental number of population 1 ($N_1$) and/or that of population 2 ($N_2$) is known ({\bf Table~\ref{parameter}}). The performance of the estimator (accuracy and precision) was quantitatively evaluated by running an individual-based model ({\bf Figure~\ref{violins}}; see also {\bf Table S1} in {\bf Supporting Information}). Our modeling framework utilizes several types of reproductive variations (survived offspring number per parent), including variance of reproductive potential within and between populations, taking account of several situations, including a body-size structure or environmental heterogeneity for reproductive success.

We presented mainly three novel points in this paper. First, we formulated probabilities of kinship pairs randomly chosen between two populations ($\pi_{\rm HS,bet}$ and $\pi_{\rm PO,bet}$), which provide estimators of the migration number or rate. While similar derivations may be found in (nongenetical) the MR method, these are limited to the situation such that sampling can be non-invasive and adult individuals are sampled. Our proposed method can avoid these limitations, where such an advantages is characterized by CKMR method for estimating population sizes. Second, we revealed that the probabilities of kinship pairs are approximately independent of reproductive potential and thus information about it is not needed for the estimation, provided that migration assumes to occur independently of the reproductive potential. This is a useful property of the estimators because it is natural that heterogeneity of reproductive potential within and between populations exits in iteroprous species but its observation is generally difficult except a body size. Third, we demonstrated applicability that the estimation can be obtained only from the genetic data (i.e., ${\hat M_5}$ and ${\hat m_4}$). There are several advantages in using the proposed estimator instead of separately estimating $\pi_{\rm HS,bet}$ (and/or $\pi_{\rm PO,bet}$) (via genetic method) and $N_1$ (and/or $N_2$) (via nongenetical method), including simplified sampling and analyzing designs and availability of unified framework of genetic analyzes for detecting HS and PO pairs (similar discussion is found for estimating the ratio of effective breeding size to the census size, $N_{\rm b}/N$, in \cite{Akita_2020}).

>>>>

\subsection{PO pair-based model}

In the folowing, we consider the probability that offspring sampled in population 1 and a parent sampled in population 2 share a PO relationship, denoted by $\pi_{\rm PO, bet}$. $\pi_{\rm PO, bet}$ can be partitioned into the above mentioned probabilities. Note that probabilities (i) and (iii) are the same as that introduced in the previous subsection, i.e., $2M/N_{1}$ and $1/M$. In addition, probability (ii) is $M/N_{2}$ by definition. Taken together, we obtain the following:
\begin{align}
\pi_{\rm PO,bet} &= \frac{2M}{N_{1}} \frac{M}{N_{2}} \frac{1}{M} \nonumber\\
&= \frac{2M}{N_{1}N_{2}}.
\label{PO-1}
\end{align}
This form also holds under a flexible setting for the reproductive potential, as noted in the {\bf Appendix 2}. Here, if all $N_1$ parents safely move to population 2 (i.e., $M = N_1$), $\pi_{\rm PO,bet}$ equals $2/N_2$. In addition, if population 2 comprises only migrants (i.e., $M = N_2$), $\pi_{\rm PO,bet}$ equals $2/N_1$. In extreme cases, these probabilities take similar forms to the PO probability that is sampled randomly within a population \cite[]{bravington2016close}.

Let $H_{\rm PO,bet}$ be the number of PO pairs found in offspring samples of size $n_{\rm O, 1}$ and parent samples of size $n_{\rm P, 2}$. Under the assumption that the total number of PO pairs in the two populations is much greater than $H_{\rm PO,bet}$, the distribution is approximated by a binomial form (i.e., $H_{\rm PO,bet} \sim {\rm Binom}[\pi_{\rm PO,bet}, n_{\rm O, 1} n_{\rm P, 2}]$). Thus, the theoretical expectation of $H_{\rm PO,bet}$ is expressed as follows: 
\begin{align}
\mathbb{E}[H_{\rm PO, bet}] &= \pi_{\rm PO,bet} n_{\rm O, 1} n_{\rm P, 2} \nonumber\\
&= \frac{ 2 n_{\rm O, 1} n_{\rm P, 2} M } { N_{1} N_{2} } \label{PO-2-1} \\
&= \frac{ 2 n_{\rm O, 1} n_{\rm P, 2} m } { N_{2} }.
\label{PO-2-2}
\end{align}
The observed number of PO pairs in a sample is defined by $\widetilde{H}_{\rm PO, bet}$, and $\mathbb{E}[H_{\rm PO, bet}]$ in Equations \ref{PO-2-1} and \ref{PO-2-2} are replaced by $\widetilde{H}_{\rm PO, bet}$, which generates the linear estimators of $M$ and $m$ as follows:
\begin{align}
\widehat{M_2} &= \frac{N_{1} N_{2} \widetilde{H}_{\rm PO,bet} } {2 n_{\rm O, 1} n_{\rm P, 2} },
\label{PO-3}
\end{align}
and
\begin{align}
\widehat{m_2} &= \frac{N_{2} \widetilde{H}_{\rm PO,bet} } {2 n_{\rm O, 1} n_{\rm P, 2} }.
\label{PO-4}
\end{align}

@@

The proposed estimators are based on the observed number of kinship pairs. Their expected number is linearly determined by the number of pairwise comparison (Eqs.~\ref{HS-2-1}-\ref{HS-2-2} and \ref{PO-2-1}-\ref{PO-2-2}), which provides sample size guidance to ensure the condition that at least one or more kinship pairs can be found. The conditions are given as follows: 
\begin{align}
n_{\rm O, 1} n_{\rm O, 2} &> \frac{ N_{2} } { 4m },
\label{RS-1}
\end{align}
and
\begin{align}
n_{\rm O, 1} n_{\rm P, 2} &> \frac{ N_{2} } { 2m }.
\label{RS-2}
\end{align}
Generally, the required sample size to ensure the above conditions is $n>\sqrt{N_2}$, where $n = n_{\rm O, 1}=n_{\rm O, 2}=n_{\rm P, 2}$. Noted that a very small $m$ value, which dramatically increases the required sample size, produces significant genetic differences between the two populations. In such cases, there are several methods to estimate migration rate using population genetics technique. Alternatively, we focus on moderate/large migration rate, which yields little genetic differences between the two populations.

Other guidance for the required sample size is obtained by an approximate lower bound on the coefficient of variation (CV) of $M_1$ or $m_1$, which is defined as $1/\sqrt{H_{\rm HS, bet}}$. This is applied in the context of classic MR \cite[]{seber1982estimation} or CKMR \cite[]{bravington2016close}. For example, to achieve a 30\% CV, the target of $H_{\rm HS, bet}$ is greater than 10, thereby providing the required sample size when the population parameters are given (e.g., $N_1$, $N_2$, and $M$). Similarly, the CV of $M_2$ or $m_2$, defined by $1/\sqrt{H_{\rm PO, bet}}$, provides the sample size required to estimate $M_2$ or $m_2$.

@@

When both HS and PO pairs, which are found between populations, are available simultaneously, we can obtain efficient estimators by combined $\widehat{M_1}$ and $\widehat{M_2}$ for migration number: 
\begin{align}
\widehat{M_3} &= \frac{ N_{1} N_{2} \left(\widetilde{H}_{\rm HS,bet} + \widetilde{H}_{\rm PO,bet}\right) } {2n_{\rm O, 1} \left(2 n_{\rm O, 2} + n_{\rm P, 2}\right) },
\label{AP-1}
\end{align}
and by combining $\widehat{m_1}$ and $\widehat{m_2}$ for the migration rate: 
\begin{align}
\widehat{m_3} &= \frac{ N_{2} \left(\widetilde{H}_{\rm HS,bet} + \widetilde{H}_{\rm PO,bet}\right) } {2n_{\rm O, 1} \left(2 n_{\rm O, 2} + n_{\rm P, 2}\right)}.
\label{AP-2}
\end{align}
For those estimators, $\widetilde{H}_{\rm HS,bet}$ and $\widetilde{H}_{\rm PO,bet}$ are weighted by sample size $n_{\rm O, 2}$ and $n_{\rm P, 2}$, respectively.

\subsection{Estimation of parent number by PO pairs in a population}

To this point, we have stated that the formulation of the estimators, presented in Equations \ref{HS-3}-\ref{HS-4} and \ref{PO-3}-\ref{PO-4}, is a function with parent numbers for each population ($N_1$ and/or $N_2$). In other words, such estimators are available only when the parent number or numbers are known. Here, we describe how we estimate unknown parent numbers by additionally using PO pairs to produce estimators for both migration number and rate that can be obtained from only the genetic data.

When PO pairs found in offspring samples and parent samples from population 2 in the second year are available, the standard parental number estimator can be obtained \cite[]{bravington2016close} as follows: 
\begin{align}
\widehat{N_2} &= \frac{ 2 n_{\rm O, 2} n_{\rm P, 2}  }{\widetilde{H}_{\rm PO, 2} + 1} ,
\label{AP-3}
\end{align}
where $\widetilde{H}_{\rm PO, 2}$ is the observed number of PO pairs found in the offspring samples of size $n_{\rm O, 2}$ and parent samples of size $n_{\rm P, 2}$. This term ``$+1$'' reduces bias, especially when $\widetilde{H}_{\rm PO, 2}$ is small \cite[e.g.,][]{ecolevol2021p}, where a similar derivation of this bias correction is provided in the literature \cite{Akita_2019}. By replacing $N_2$ by $\widehat{N_2}$ in Eqs.~\ref{AP-1} and \ref{AP-2}, we obtain the following estimators: 
\begin{align}
\widehat{M_4} &=  \frac{ n_{\rm O, 2} n_{\rm P, 2} N_{1} \left(\widetilde{H}_{\rm HS,bet} + \widetilde{H}_{\rm PO,bet}\right) } {n_{\rm O, 1} \left(2 n_{\rm O, 2} + n_{\rm P, 2}\right) \left(\widetilde{H}_{\rm PO, 2} + 1\right)},
\label{AP-4}
\end{align}
and
\begin{align}
\widehat{m_4} &=  \frac{ n_{\rm O, 2} n_{\rm P, 2}\left(\widetilde{H}_{\rm HS,bet} + \widetilde{H}_{\rm PO,bet}\right)} {n_{\rm O, 1} \left(2 n_{\rm O, 2} + n_{\rm P, 2}\right) \left(\widetilde{H}_{\rm PO, 2} + 1\right)}.
\label{AP-5}
\end{align}
While the estimator of the migration number (Equation~\ref{AP-4}) requires the (unknown) parent number in population 1 ($N_{1}$), the estimator of the migration rate (Equation~\ref{AP-5}) can be obtained using only the observed number of HS and PO pairs.

Similar to estimating $N_2$, when PO pairs found in offspring samples and parent samples from population 1 in the first year are available, we can obtain the estimator of $N_1$ as follows:
\begin{align}
\widehat{N_1} &= \frac{ 2 n_{\rm O, 1} n_{\rm P, 1} }{\widetilde{H}_{\rm PO, 1} +1},
\label{AP-6}
\end{align}
where $\widetilde{H}_{\rm PO, 1}$ is the observed number of PO pairs in the offspring samples of size $n_{\rm O, 1}$ and parent samples of size $n_{\rm P, 1}$. By replacing $N_1$ with $\widehat{N_1}$ in Eq.~\ref{AP-4}, we obtain a migration number estimator that can be calculated from only genetic data, which is expressed as follows:
\begin{align}
\widehat{M_5} &=  \frac{ 2 n_{\rm O, 1} n_{\rm P, 1} n_{\rm O, 2} n_{\rm P, 2} \left(\widetilde{H}_{\rm HS,bet} + \widetilde{H}_{\rm PO,bet}\right)} {n_{\rm O, 1} \left(2 n_{\rm O, 2} + n_{\rm P, 2}\right) \left(\widetilde{H}_{\rm PO, 1} + 1\right) \left(\widetilde{H}_{\rm PO, 2} + 1\right)}.
\label{AP-7}
\end{align}

Table \ref{parameter} summarizes the conditions for estimators in terms of whether $N_1$ and/or $N_2$ are known and which type of samples are required for estimation. In addition, Table \ref{parameter} identifies the kinship type required to calculate the estimator.

@@

The proposed estimators are derived under the assumption that reproductive potential is equal among individuals. As described in {\bf Appendix 2}, this assumption can be relaxed, and the estimators still hold, which is exemplified in iteroparous species that may show significant variation in reproductive potential among individuals. In addition, the estimators hold for most cases where the mean reproductive potential differes between populations 1 and 2, e.g., when the environmental condition for reproductive success varies between populations. Here, the required condition is that the migration event does not depend on the degree of reproductive potential (refer to {\bf Appendix 3} for additional details).

@@

We developed an individual-based model that tracks kinship relationships to evaluate the performance of the estimators. Here, the population structure was assumed to be identical to that in the development of the estimators. Population 1 and 2 comprised $N_1$ and $N_2$ parents with an equal sex ratio, and their offspring number was assumed to follow the geometric distribution with mean ${\bar \lambda_1}$ and ${\bar \lambda_2}$ (i.e., Poisson reproduction with mean $\lambda_{i,1}$ and $\lambda_{j,2}$, which follows the exponential distribution with mean ${\bar \lambda_1}$ and ${\bar \lambda_2}$), respectively. Migrant parents were selected randomly from population 1 at the end of the first year, and each offspring retained IDs of the parents, thereby making it possible to trace an HS and PO relationship.

Here, the parameter set ($N_1$, $N_2$, $M$, ${\bar \lambda_1}$, ${\bar \lambda_2}$, $n_{\rm O,1}$, $n_{\rm O,2}$, $n_{\rm P,1}$, $n_{\rm P,2}$) is given. We simulated a population history and a sampling process, which generates proposed estimators. Note that this process was repeated 1000 times, which allows us to construct the distribution of the estimators for each parameter set. All scripts (written by C{++} and R languages) and documentation for these analyses are available at https://github.com/teTUNAakita/CKMRmig.

@@

We evaluated the performance of ${\hat M}$s or ${\hat m}$s numerically for a case with variable reproductive potential among parents. Here, the scaled statistical properties of ${\hat m_1}$,  ${\hat m_2}$,  ${\hat m_3}$ and ${\hat m_4}$ were completely the same as ${\hat M_1}$, ${\hat M_2}$, ${\hat M_3}$, and ${\hat M_4}$; thus, hereafter, we only demonstrate the results for ${\hat M_1}$-${\hat M_5}$. {\bf Figure~\ref{violins}} shows the distribution of the relative bias of ${\hat M}$ for limiting cases where parent and offspring sample numbers are identical (i.e., $n=n_{\rm P,1}=n_{\rm P,2}=n_{\rm O,1}= n_{\rm O,2}$) and parent numbers in the two populations are also identical (i.e., $N=N_1=N_2$). The relative bias is calculated by applying the outputs of the individual-based model, which is defined as ``(averaged estimator $-$ true value)/true value.'' Refer to {\bf Table S1} in {\bf Supporting Information} for a full list of the parameter sets used to evaluate performance (relative bias and CV of ${\hat M}$s).

@@

First, we evaluated the accuracy of ${\hat M}$s based on the relative bias. As expected, for most of the investigated parameter sets, we observed that their relative bias was less than 5\%, as shown in {\bf Table S1} in {\bf Supporting Information}. Thus, when the assumption that kinships are detected without any error is satisfied, it is reasonable to consider ${\hat M}$s nearly unbiased estimators.

Next, we evaluated the precision of ${\hat M}$s based on the CV value. {\bf Table S1} in {\bf Supporting Information} shows the CV value, and the violin plot in {\bf Figure~\ref{violins}} visualizes the degree of precision. For each estimator ${\hat M}$, we found that precision increases with increasing sample size. Note that the total sample size depends on the estimator. For example, $n=50$ in Fig.~\ref{violins}, the total sample sizes of ${\hat M_1}$, ${\hat M_2}$, ${\hat M_3}$, ${\hat M_4}$, and ${\hat M_5}$ were 100 ($=n_{\rm O,1}+n_{\rm O,2}$), 100 ($=n_{\rm O,1}+n_{\rm P,2}$), 150 ($=n_{\rm O,1}+n_{\rm O,2}+n_{\rm P,2}$), 150 ($=n_{\rm O,1}+n_{\rm O,2}+n_{\rm P,2}$), and 200 ($=n_{\rm O,1}+n_{\rm O,2}+n_{\rm P,1}+n_{\rm P,2}$), respectively.

As the sample size or number of migrants increase, precision increases, and the shape of the distribution asymptotically becomes symmetric ({\bf Fig.~\ref{violins}}) because an increasing sample size or increasing number of migrants is likely to increase the observed number of kinship pairs found between the two populations ($\widetilde{H}_{\rm HS,bet}$ or $\widetilde{H}_{\rm PO,bet}$) and decreases the variance of those kinship pair numbers. In addition, the number of PO pairs in a given population ($\widetilde{H}_{\rm PO,1}$ or $\widetilde{H}_{\rm PO,2}$) contributes to the precision of estimating $N_1$ or $N_1$, respectively, thereby providing ${\hat M_4}$ and ${\hat M_5}$ with higher precision.

Note that our simulation can handle case with and without invasive sampling. Invasive sampling potentially affects the level of $m$ because the sampled parents from population 1 have no chance to move to population 2; thus, such parents cannot be sampled. In this case, $m$ should be defined by $M/(N_1-n_{P,1})$ rather than $M/N_1$, although this does not affect the estimator ${\hat M}$s, as demonstrated in {\bf Table S1} ({\bf Supporting Information}).

Finally, we investigated the case where the parental number in population 1 is much greater in population 2 (i.e., $N_1 \gg N_2$); thus, migrants represent a  large proportion of $N_2$. For such cases, e.g., $(N_1,N_2,M)=(10^4,10^3,500)$, we confirmed the similar property of ${\hat M}$s to the $N_1 = N_2$ situation, as shown in {\bf Table S1} ({\bf Supporting Information}), which suggests the robustness of ${\hat M}$s in the presence of heterogeneous population sizes.

@@

In this study, we theoretically developed estimators for the contemporary migration number (${\hat M_1}$--${\hat M_5}$) and migration rate (${\hat m_1}$--${\hat m_4}$) of parents between two predefined populations in iteroparous species. The proposed estimators are based on the known PO relationship and HS relationships observed between and within the two populations without any error in terms of kinship assignment. Users can select the appropriate estimator for cases where the parental number of population 1 ($N_1$) and/or that of population 2 ($N_2$) is known ({\bf Table~\ref{parameter}}). The performance of the estimator (accuracy and precision) was evaluated quantitatively by performing an individual-based simulation ({\bf Fig.~\ref{violins}} and {\bf Table S1} in {\bf Supporting Information}). The proposed modeling framework utilizes several types of reproductive variations (i.e., the number of survived offspring per parent), including the variance of reproductive potential within and between populations, in consideration of several situations, including a body-size structure or environmental heterogeneity for reproductive success.

Our primary contributions are summarized as follows. First, we formulated the probabilities of kinship pairs randomly selected between two populations ($\pi_{\rm HS,bet}$ and $\pi_{\rm PO,bet}$), which provides migration number or rate estimators. While similar derivations may be found in the (nongenetic) MR method, these are limited to cases where sampling is non-invasive and adult individuals are sampled. The proposed method can avoid these limitations, where such an advantages is characterized by the CKMR method for estimating population sizes. Second, we have demonstrated that the probabilities of kinship pairs are approximately independent of the reproductive potential; thus, information about it is not required to realize the estimation provided that migration is assumed to occur independently of the reproductive potential. This is a useful property of the estimators because it is natural that heterogeneity of reproductive potential within and between populations exits in iteroparous species but its observation is generally difficult. Third, we have demonstrated that estimators can be obtained using only genetic data (i.e., ${\hat M_5}$ and ${\hat m_4}$). There are several advantages in using the proposed estimator rather than estimating $\pi_{\rm HS,bet}$ (and/or $\pi_{\rm PO,bet}$) separately (via a genetic method) and $N_1$ (and/or $N_2$) (via a nongenetic method), including simplified sampling processes and analyzing designs and the availability of a unified framework of genetic analyses for detecting HS and PO pairs (similar discussion is found for estimating the ratio of effective breeding size to the census size, $N_{\rm b}/N$, in \cite{Akita_2020}).
